---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# usdaplants

<!-- badges: start -->
<!-- badges: end -->

usdaplants is a convenience data package providing the USDA PLANTS database as
tibbles. It includes plant taxonomy, growth characteristics, and state-level
distribution data for over 32,000 plant species in the United States.

## Installation

You can install the development version of usdaplants from
[GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("kjhealy/usdaplants")
```

## Datasets

The package provides three datasets:

- **`plants`** -- 32,800 plants with taxonomy, duration, and growth habit.
- **`plant_characteristics`** -- Growth and habitat characteristics for 2,162 species.
- **`plant_distribution`** -- Native/introduced status by state (241,601 rows).

All three datasets are linked by the `symbol` column (the USDA PLANTS identifier).

## Examples

```{r example}
library(usdaplants)

plants
```

```{r}
# Filter to trees
plants[grepl("Tree", plants$growth_habit), ]
```

```{r}
# Characteristics for drought-tolerant plants
plant_characteristics[plant_characteristics$drought_tolerance == "High", c("symbol", "drought_tolerance", "height_mature_feet")]
```

```{r}
# Plants native to New York
ny_native <- plant_distribution[
  plant_distribution$state == "New York" & plant_distribution$status == "native",
]
ny_native
```
